### 2017-02-28 Doc/ms.Rmd
update with bookdown package

---
title: "Your Awesome Title"
author: "Author One and Author Two"
date: '`r format(Sys.time())`'
output:
  bookdown::pdf_document2:
    includes:
      before_body: doc_prefix.tex
      in_header: preamble.tex
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
    toc: no
  bookdown::html_document2:
    number_sections: no
    theme: readable
    toc: yes
  bookdown::tufte_html2:
    number_sections: no
    toc: yes
  bookdown::word_document2: null
fontsize: 12pt
link-citations: yes
csl: oikos.csl
bibliography: ref.bib
---

```{r knitr_options, echo=FALSE}
library(knitr)
# rstudio will set the folder where .Rmd file seats as work directory
# set it back to the folder where .Rproj seats
opts_knit$set(root.dir = normalizePath("../")) 
opts_chunk$set(fig.align='center', cache=F, warning = FALSE, message = T, echo = FALSE)
options(digits = 3, width = 88)
```

**Running headline**: Environment and species richness

**Abstract**: Your awesome abstract here.

\clearpage

# Introduction

Here is your introduction. It should describe clearly the rationale for the study being done and the previous work related with the study. It should also tell readers about your specific hypothese/questions being addressed. Citations will be like this [@adair_single-pool_2010], or [e.g., @clark_loss_2008], or [@eriksson_seed_1993; @williamson_dissolved_1999]

Here is the second paragraph of the introduction.

# Methods

Here is the method section. You can include equations easily. For inline equations, use $\text{var}(X) = p(1-p)$. For display equation, use

$$\text{var}(X) = p(1-p)$$

## Results

#### Tables

Insert tables by `kable` in knitr package in R. Then cross-reference it back with: see Table \@ref(tab:tableName).

```{r tableName,results='asis',echo=FALSE, message=F, warning=F}
source("R/rcode.R")
knitr::kable(sprich, booktabs = T, caption = "Caption here.")
```

Put results inline, e.g. the mean species richness is `r meansprich`.

#### Insert tables by `xtable` package in R

Show as Table. \ref{t:anova}:

```{r echo=F,results='asis',warning=F,message=F}
library(xtable)
envi.summ = xtable(summary(envi.aov),label = "t:anova", caption = "Caption here")
print(envi.summ,
		caption.placement = "top",
	  floating=T,
	  hline.after=NULL,
	  comment = FALSE,
	  add.to.row=list(pos=list(-1,0, nrow(envi.summ)),
	  command=c(
		   '\\toprule\n',
		  '\\midrule\n',
		   '\\bottomrule\n')))
```

#### Insert tables by hand

Show as Table. \@ref(tab:byhand):

Table: (#tab:byhand) Caption here.

Col A             Col B       Col C                   Col D
------------      -------     ------------------      ------------------
row 1             190         $112 \pm 2$             $233 \pm 3$
$\eta$            0.13        0.12                    0.12
$\eta^2$          0.14        0.13                    0.50
$\eta^3$          0.15        0.31                    0.52

#### Figures

Insert figure by code chunk. And cross-ref it back as Figure \@ref(fig:figName).

```{r figName, fig.width=7, fig.asp=1, fig.cap="Your caption here."}
ggplot(data = envi.sprich, aes(x = value, y = sprich)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_wrap(~envi, scales = "free_x") + theme_bw()
```

Or if you already have the figure: 
And cite it as Figure \@ref(fig:fig2).

```{r fig2, out.width='70%', fig.cap="Caption here."}
knitr::include_graphics(path = paste0(normalizePath(getwd()), "/Figs/plot.pdf"))
```

\clearpage

# References