### 2021-11-02 Doc/respond_to_reviewer.Rmd
update demo

---
title: "Response to reviewers"
author: "Daijiang Li"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    highlight: zenburn
    keep_tex: no
    template: template/respond_to_reviewer.tex
    toc: yes
    toc_depth: 3
    latex_engine: xelatex
  html_document:
    css: template/respond_to_reviewer.css
    keep_md: no
    theme: readable
    toc: yes
  word_document: default
---

```{r knitr_options, echo=FALSE}
library(knitr)
# rstudio will set the folder where .Rmd file seats as work directory
# set it back to the folder where .Rproj seats
opts_knit$set(root.dir = normalizePath("../")) 
opts_chunk$set(fig.align = 'center', cache = FALSE, warning = FALSE,
  message = TRUE, echo = FALSE, fig.pos = "H", out.extra = "")
options(digits = 3, width = 88,
        # use pdf file automatically when compile to pdf
        knitr.graphics.auto_pdf = TRUE, 
        knitr.kable.NA = '')

# download template files if not available
tpl_1 = 'https://raw.githubusercontent.com/daijiang/workflow_demo/master/Doc/template/respond_to_reviewer.tex'
tpl_2 = 'https://raw.githubusercontent.com/daijiang/workflow_demo/master/Doc/template/respond_to_reviewer.css'
# change directory accordingly
if(!file.exists(tpl_1f <- 'template/respond_to_reviewer.tex')) download.file(tpl_1, tpl_1f)
if(!file.exists(tpl_2f <- 'template/respond_to_reviewer.css')) download.file(tpl_2, tpl_2f)

if(knitr::is_latex_output() | knitr::is_html_output()){
  library(kableExtra)
} else {
  options(kableExtra.auto_format = FALSE) # for docx
}
```

>09-Nov-2017

>Dear Dr. Li

>We have now completed the review process on your manuscript ....

>You will find the reports of the referees (if any), and the comments of the handling editor, ...

>I invite you to

>Thank you again for submitting to ....

>Yours sincerely

>Editor-in-chief

### EDITOR-IN-CHIEF'S COMMENTS TO AUTHORS

>xxxxxxxxxx

Dear Editors,

Thank you for your help with our manuscript. The constructive comments from reviewers and Editors have greatly improved it.

Sincerely,

Daijiang Li

### EDITOR'S COMMENTS TO AUTHORS

>Editor:   
>Comments to the Author:
>Dear authors

>Thanks for your resubmission, ...

>Best wishes,
>Editor...

We have added qualifiers in these subsections as suggested.

>L299 change ‘reveal’ to ‘suggest’ or ‘imply’ or similar.

Changed 'reveal' to "suggest."

### REVIEWER COMMENTS TO AUTHORS

>If there are no comments above, the reviewer may have chosen to upload a file. Please check ScholarOne Manuscripts for any attached files, which may be found at the end of the decision letter in your author centre under 'Manuscripts with decisions'.

Thank you and here is our responses.